---
title: "Colonist Replay Analysis"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    fig-width: 11
    fig-height: 6.5
    fig-dpi: 150
    theme: cosmo
    embed-resources: true
    include-in-header:
      text: |
        <style>
          /* Smooth fade + bounce when a new panel-tabset pane becomes active */
          .panel-tabset .tab-content .tab-pane.active {
            animation: dotsArrive 0.9s cubic-bezier(0.2, 0.8, 0.2, 1);
          }
          @keyframes dotsArrive {
            0%   { opacity: 0; transform: translateY(-18px) scale(0.96); filter: blur(4px); }
            45%  { opacity: 0.75; transform: translateY(8px) scale(1.01); filter: blur(1px); }
            75%  { opacity: 0.95; transform: translateY(-2px) scale(1.002); filter: blur(0); }
            100% { opacity: 1; transform: translateY(0) scale(1); filter: blur(0); }
          }
          /* Slight stagger for images inside tabs so dots look like they "fly in" */
          .panel-tabset .tab-pane.active .cell-output-display img {
            animation: imgPop 1.05s cubic-bezier(0.25, 1, 0.3, 1);
          }
          @keyframes imgPop {
            0%   { opacity: 0; transform: scale(0.85) translateY(10px); }
            60%  { opacity: 0.9; transform: scale(1.015) translateY(-2px); }
            100% { opacity: 1; transform: scale(1) translateY(0); }
          }
          /* Tab buttons pop a little when selected */
          .nav-tabs .nav-link.active {
            transition: transform 180ms ease, color 180ms ease;
            transform: translateY(-1px);
          }
        </style>
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(scales)
library(ggrepel)
library(ggimage)
library(gt)
library(reactable)
library(htmltools)

setwd(here::here())
source("analysis/load_replays.R")
d <- load_all()

# Colors
RESOURCE_COLORS <- c(
  lumber = "#517d19",
  brick  = "#9c4300",
  wool   = "#F2EBD8",
  grain  = "#f0ad00",
  ore    = "#7b6f83",
  desert = "#cccccc"
)
NON_RESOURCE_COLOR <- "#4fa6eb"   # for dice, activity, anything not tied to a resource
HIGHLIGHT_COLOR    <- "#9c4300"   # for callouts (e.g. 7s, red numbers)

RESOURCE_LEVELS <- c("lumber", "brick", "wool", "grain", "ore")

RESOURCE_IMAGES <- c(
  lumber = here::here("images", "lumber.png"),
  brick  = here::here("images", "brick.png"),
  wool   = here::here("images", "sheep.png"),
  grain  = here::here("images", "wheat.png"),
  ore    = here::here("images", "ore.png")
)

BAR_BORDER   <- "black"
BAR_LW       <- 0.8
AXIS_LW      <- 1.6
POINT_STROKE <- 1.1

theme_colonist <- function(base_size = 13) {
  theme_minimal(base_size = base_size, base_family = "Helvetica") +
    theme(
      text               = element_text(face = "bold", color = "gray12"),
      plot.title         = element_text(face = "bold", size = rel(1.4), margin = margin(b = 4)),
      plot.subtitle      = element_text(face = "bold", color = "gray30", margin = margin(b = 14)),
      plot.caption       = element_text(face = "plain", color = "gray55",
                                        size = rel(0.78), margin = margin(t = 10)),
      plot.title.position   = "plot",
      plot.caption.position = "plot",
      axis.title         = element_text(face = "bold", color = "gray15"),
      axis.title.x       = element_text(margin = margin(t = 10)),
      axis.title.y       = element_text(margin = margin(r = 10)),
      axis.text          = element_text(face = "bold", color = "gray20"),
      axis.line.x        = element_line(color = "black", linewidth = AXIS_LW),
      axis.line.y        = element_line(color = "black", linewidth = AXIS_LW),
      axis.ticks         = element_line(color = "black", linewidth = 1),
      axis.ticks.length  = unit(5, "pt"),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = "gray90", linewidth = 0.4),
      strip.text         = element_text(face = "bold", size = rel(1.0)),
      legend.position    = "top",
      legend.title       = element_text(face = "bold"),
      legend.text        = element_text(face = "bold"),
      legend.key.width   = unit(18, "pt")
    )
}
```

```{r headline}
n_games <- nrow(d$games)
n_players <- length(unique(d$rosters$username))
```

*Based on **`r n_games` games** and **`r n_players` players** in `replays/`.*

## Player Data

Switch between totals across all games and per-game averages. Every column is sortable — click any header.

```{r player-data-prep}
# Per-game sums of cards lost to 7 rolls (from seven_discards event table)
seven_losses <- d$seven_discards |>
  group_by(game_id, player_color) |>
  summarise(cards_lost_to_7 = sum(n_cards, na.rm = TRUE), .groups = "drop")

player_totals <- d$player_stats |>
  left_join(seven_losses, by = c("game_id", "player_color")) |>
  mutate(cards_lost_to_7 = coalesce(cards_lost_to_7, 0L)) |>
  group_by(username) |>
  summarise(
    games                = n(),
    dev_cards_bought     = sum(dev_cards_bought),
    dev_cards_played     = sum(dev_cards_used),
    roads                = sum(built_road),
    settlements          = sum(built_settlement, na.rm = TRUE),
    cities               = sum(built_city, na.rm = TRUE),
    trades_proposed      = sum(proposed_trades),
    trades_accepted      = sum(successful_trades),
    bank_trades          = sum(n_bank_trades),
    times_robbed_others  = sum(n_steals_done),
    resources_blocked    = sum(resource_income_blocked),
    cards_lost_to_7      = sum(cards_lost_to_7),
    resources_used       = sum(resources_used),
    cards_from_rolls     = sum(gained_total),
    .groups = "drop"
  ) |>
  mutate(trade_completion_rate = ifelse(trades_proposed > 0,
                                        trades_accepted / trades_proposed,
                                        NA_real_)) |>
  relocate(trade_completion_rate, .after = trades_accepted) |>
  arrange(desc(resources_used))

player_per_game <- player_totals |>
  mutate(across(
    -c(username, games, trade_completion_rate),
    ~ round(.x / games, 1)
  ))

metric_labels <- c(
  username              = "Player",
  games                 = "Games",
  dev_cards_bought      = "Dev Cards Bought",
  dev_cards_played      = "Dev Cards Played",
  roads                 = "Roads",
  settlements           = "Settlements",
  cities                = "Cities",
  trades_proposed       = "Trades Proposed",
  trades_accepted       = "Trades Accepted",
  trade_completion_rate = "Trade Completion Rate",
  bank_trades           = "Bank Trades",
  times_robbed_others   = "Times Robbed Others",
  resources_blocked     = "Resources Blocked by Robber",
  cards_lost_to_7       = "Cards Lost to 7",
  resources_used        = "Resources Used",
  cards_from_rolls      = "Cards From Rolls"
)

make_player_table <- function(df, grey_one_game = FALSE) {
  col_defs <- lapply(names(metric_labels), function(col) {
    if (col == "username") {
      colDef(name = metric_labels[[col]], minWidth = 120, sticky = "left",
             style = list(fontWeight = "bold"))
    } else if (col == "games") {
      colDef(name = metric_labels[[col]], align = "center", minWidth = 70)
    } else if (col == "trade_completion_rate") {
      colDef(
        name = metric_labels[[col]],
        align = "center", minWidth = 110,
        format = colFormat(percent = TRUE, digits = 0),
        style = function(value, index) {
          if (grey_one_game && df$games[index] <= 1) {
            return(list(background = "#d9d9d9", color = "#666", fontStyle = "italic"))
          }
          if (is.na(value)) return(NULL)
          list(background = sprintf("rgba(81, 125, 25, %s)",
                                    round(min(value, 1) * 0.45, 2)))
        }
      )
    } else {
      colDef(
        name = metric_labels[[col]],
        align = "center", minWidth = 95,
        style = function(value, index) {
          if (grey_one_game && df$games[index] <= 1) {
            return(list(background = "#d9d9d9", color = "#666", fontStyle = "italic"))
          }
          max_val <- max(df[[col]][df$games > 1], na.rm = TRUE)
          if (!is.finite(max_val) || max_val <= 0) return(NULL)
          frac <- min(value / max_val, 1)
          list(background = sprintf("rgba(81, 125, 25, %s)", round(frac * 0.45, 2)))
        }
      )
    }
  })
  names(col_defs) <- names(metric_labels)
  reactable(
    df,
    columns = col_defs,
    defaultSorted = list(resources_used = "desc"),
    sortable = TRUE,
    striped = TRUE,
    compact = TRUE,
    highlight = TRUE,
    bordered = TRUE,
    defaultPageSize = 20,
    theme = reactableTheme(
      headerStyle = list(fontWeight = "bold", background = "#f7f7f7"),
      style = list(fontFamily = "Helvetica")
    )
  )
}
```

::: {.panel-tabset}

## Total

```{r player-table-total}
make_player_table(player_totals, grey_one_game = FALSE)
```

## Per Game

```{r player-table-per-game}
make_player_table(player_per_game, grey_one_game = TRUE)
```

:::

**Column meanings**

- **Trades Proposed** = trade offers this player put out
- **Trades Accepted** = offers this player **proposed that someone else accepted** (NOT trades where they were on the accepting side)
- **Trade Completion Rate** = Trades Accepted / Trades Proposed (percentage of their offers that closed)
- **Bank Trades** = 4:1 / 3:1 / 2:1 port trades with the bank
- **Times Robbed Others** = times this player placed the robber and pulled a card from an opponent
- **Resources Blocked by Robber** = times one of this player's tiles would have produced, but the robber was on it
- **Cards Lost to 7** = total cards discarded on 7-rolls (when hand exceeded the discard limit)
- **Resources Used** = cards spent on builds, dev cards, and trades (Colonist's own `resourcesUsed`)
- **Cards From Rolls** = total resource cards received from dice rolls
- Players with only **1 game** show greyed cells in the Per Game view since the average is just the single-game total

## Overall average price per resource

"Going rate" for each resource, pooled across every completed player-to-player trade. Only trades where the requested side was a single uniform resource are counted.

```{r price-overall, fig.height=5}
p2p_prices <- d$trades_accepted |>
  mutate(
    n_given = lengths(given),
    n_received = lengths(received),
    received_uniform = map_chr(received_names, ~ {
      if (length(.x) == 0 || length(unique(.x)) > 1) NA_character_ else .x[1]
    }),
    price = n_given / n_received
  ) |>
  filter(!is.na(received_uniform), price > 0)

overall_prices <- p2p_prices |>
  group_by(resource = received_uniform) |>
  summarise(avg_price = mean(price), n_trades = n(), .groups = "drop") |>
  mutate(
    resource = factor(resource, levels = RESOURCE_LEVELS),
    image = RESOURCE_IMAGES[as.character(resource)]
  ) |>
  arrange(avg_price)

ymax <- max(overall_prices$avg_price)

ggplot(overall_prices,
       aes(x = reorder(resource, avg_price), y = avg_price)) +
  # place image on the left of each bar (axis side), outside bar area
  geom_col(aes(fill = resource),
           width = 0.65, color = BAR_BORDER, linewidth = BAR_LW) +
  geom_image(aes(image = image, y = -ymax * 0.08),
             size = 0.11, by = "width", asp = 2.0) +
  geom_text(aes(label = sprintf("%.2f (n=%d)", avg_price, n_trades)),
            hjust = -0.15, color = "gray15",
            size = 4.2, fontface = "bold") +
  coord_flip(clip = "off") +
  scale_fill_manual(values = RESOURCE_COLORS, guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0.12, 0.25)),
                     limits = c(-ymax * 0.10, NA)) +
  labs(title = "Overall average price per resource",
       subtitle = "Pooled across every completed P2P trade",
       x = NULL, y = "Cards given per card received") +
  theme_colonist() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(10, 20, 10, 40)
  )
```

## Where players put their initial settlements

### Resources adjacent to initial settlements

Use the tabs to switch between the combined view and individual players. How often each resource touches a 1st vs 2nd starting settlement.

```{r initial-resources-setup}
initial_tiles_named <- d$settlement_tiles |>
  filter(is_initial, resource != "desert") |>
  left_join(d$rosters |> select(game_id, player_color, username),
            by = c("game_id", "player_color"))

plot_initial_resources <- function(data, title_suffix) {
  if (nrow(data) == 0) {
    return(ggplot() + theme_void() +
             labs(title = paste0("Resource adjacency -- ", title_suffix),
                  subtitle = "No data"))
  }
  d2 <- data |>
    count(placement_order, resource, name = "count") |>
    group_by(placement_order) |>
    mutate(
      share = count / sum(count),
      resource = factor(resource, levels = RESOURCE_LEVELS),
      image = RESOURCE_IMAGES[as.character(resource)],
      placement_label = ifelse(placement_order == 1,
                               "1st settlement", "2nd settlement")
    ) |>
    ungroup()

  ymax <- max(d2$count)

  ggplot(d2, aes(x = resource, y = count, fill = resource)) +
    geom_col(width = 0.7, color = BAR_BORDER, linewidth = BAR_LW) +
    geom_text(aes(label = sprintf("%d\n(%.0f%%)", count, share * 100)),
              vjust = -0.3, size = 3.8, fontface = "bold", color = "gray15",
              lineheight = 0.9) +
    geom_image(aes(image = image, y = -ymax * 0.08),
               size = 0.11, by = "width", asp = 2.0) +
    scale_fill_manual(values = RESOURCE_COLORS, guide = "none") +
    scale_y_continuous(expand = expansion(mult = c(0.10, 0.20))) +
    facet_wrap(~ placement_label) +
    labs(
      title = paste0("Resource adjacency of initial settlements -- ", title_suffix),
      subtitle = "How often each resource touches a 1st / 2nd settlement",
      x = NULL, y = "Times adjacent"
    ) +
    theme_colonist() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      panel.grid.major.y = element_line(color = "gray90", linewidth = 0.4)
    )
}
```

```{r initial-resources-tabs, results='asis', fig.height=5.2, fig.width=10}
players_init <- sort(unique(initial_tiles_named$username))
cat("::: {.panel-tabset}\n\n")
cat("## Combined\n\n")
print(plot_initial_resources(initial_tiles_named, "all players"))
cat("\n\n")
for (p in players_init) {
  cat("## ", p, "\n\n", sep = "")
  print(plot_initial_resources(
    initial_tiles_named |> filter(username == p), p))
  cat("\n\n")
}
cat(":::\n")
```

### Dice numbers touched by initial settlements

```{r initial-dice-bar, fig.height=5}
dice_touched <- d$settlement_tiles |>
  filter(is_initial, !is.na(dice_number), dice_number > 0) |>
  count(placement_order, dice_number, name = "count") |>
  mutate(
    placement_label = ifelse(placement_order == 1,
                             "1st settlement", "2nd settlement"),
    is_red = dice_number %in% c(6L, 8L)
  )

ggplot(dice_touched,
       aes(x = factor(dice_number), y = count, fill = is_red)) +
  geom_col(width = 0.72, color = BAR_BORDER, linewidth = BAR_LW) +
  geom_text(aes(label = count), vjust = -0.4,
            size = 3.8, fontface = "bold", color = "gray15") +
  scale_fill_manual(values = c("TRUE" = "#2770a8", "FALSE" = NON_RESOURCE_COLOR),
                    guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  facet_wrap(~ placement_label) +
  labs(
    title = "Dice numbers of initial settlement tiles",
    subtitle = "How often each number is adjacent to a starting settlement (6 and 8 darker)",
    x = "Tile number", y = "Times adjacent"
  ) +
  theme_colonist()
```

### Expected production per settlement

Each settlement's "expected production" scores its adjacent tiles by how often they'd produce on a roll (6/8 = 5, 5/9 = 4, 4/10 = 3, 3/11 = 2, 2/12 = 1) and sums across adjacent tiles. Max possible = 15 (three 6/8 tiles). Shown for the 1st through 4th settlements each player places. Use the tabs to view one player at a time; the grey bar in each player tab is the overall mean for comparison.

```{r expected-prod-data}
placements_1_4 <- d$settlement_tiles |>
  filter(placement_order <= 4, !is.na(pips)) |>
  group_by(game_id, player_color, corner_id, placement_order) |>
  summarise(
    expected_prod = sum(pips, na.rm = TRUE),
    n_tiles = n(),
    .groups = "drop"
  ) |>
  mutate(placement_label = factor(
    paste0(
      placement_order,
      c("st", "nd", "rd", "th")[pmin(placement_order, 4)],
      " settlement"
    ),
    levels = c("1st settlement", "2nd settlement",
               "3rd settlement", "4th settlement")
  ))

# Join player usernames for per-player breakdown
placements_with_name <- placements_1_4 |>
  left_join(d$rosters |> select(game_id, player_color, username),
            by = c("game_id", "player_color"))

# Colors for settlement order — distinct palette, not resource-themed
SETTLEMENT_PAL <- c(
  "1st settlement" = "#3b6cb1",
  "2nd settlement" = "#d97a3e",
  "3rd settlement" = "#6aa15d",
  "4th settlement" = "#a54878"
)

# Overall means per settlement order (used as grey reference line in player tabs)
overall_means <- placements_1_4 |>
  group_by(placement_label) |>
  summarise(overall_mean = mean(expected_prod), .groups = "drop")
```

```{r expected-prod-player-fn}
plot_expected_prod <- function(data, title_suffix, show_overall = FALSE) {
  if (nrow(data) == 0) {
    return(ggplot() + theme_void() +
           labs(title = paste0("Expected production -- ", title_suffix),
                subtitle = "No data"))
  }

  # Per-group stats: mean to label, max to position the label
  group_stats <- data |>
    group_by(placement_label) |>
    summarise(
      mean_val = mean(expected_prod),
      max_val  = max(expected_prod),
      .groups = "drop"
    )

  y_ceiling <- 17  # extra room for mean label above topmost dot

  p <- ggplot(data, aes(x = placement_label, y = expected_prod))

  # Grey overall-mean bar behind the player's jitter (only in player tabs)
  if (show_overall) {
    p <- p +
      geom_crossbar(
        data = overall_means,
        aes(x = placement_label, y = overall_mean,
            ymin = overall_mean, ymax = overall_mean),
        color = "gray55", linewidth = 1.0, width = 0.55,
        inherit.aes = FALSE
      )
  }

  p +
    geom_jitter(
      aes(fill = placement_label),
      width = 0.18, height = 0.05,
      shape = 21, size = 4, stroke = POINT_STROKE,
      color = "black", alpha = 0.9
    ) +
    stat_summary(fun = mean, geom = "crossbar", width = 0.45,
                 color = "black", linewidth = 1.0, fill = NA) +
    geom_text(
      data = group_stats,
      aes(x = placement_label, y = max_val + 1.5,
          label = sprintf("mean = %.1f", mean_val)),
      fontface = "bold", color = "black", size = 4.4,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = SETTLEMENT_PAL, guide = "none") +
    scale_y_continuous(breaks = seq(0, 15, 3),
                       limits = c(0, y_ceiling)) +
    labs(
      title = paste0("Expected production per settlement -- ", title_suffix),
      subtitle = "Sum across adjacent tiles (bar = mean; grey bar in player tabs = overall mean)",
      x = NULL, y = "Expected production"
    ) +
    theme_colonist() +
    theme(axis.text.x = element_text(size = rel(1.05)))
}
```

```{r expected-prod-player-tabs, results='asis', fig.height=4.8, fig.width=9.5}
players <- sort(unique(placements_with_name$username))
cat("::: {.panel-tabset}\n\n")

cat("## Combined\n\n")
print(plot_expected_prod(placements_with_name, "all players",
                         show_overall = FALSE))
cat("\n\n")

for (p in players) {
  cat("## ", p, "\n\n", sep = "")
  print(plot_expected_prod(
    placements_with_name |> filter(username == p),
    p,
    show_overall = TRUE
  ))
  cat("\n\n")
}
cat(":::\n")
```

## Dice distribution (pooled)

```{r dice-pooled, fig.height=4.8}
pooled_dice <- d$dice_histogram |>
  group_by(total) |>
  summarise(count = sum(count), .groups = "drop") |>
  mutate(is_seven = total == 7)

ggplot(pooled_dice, aes(x = factor(total), y = count, fill = is_seven)) +
  geom_col(width = 0.78, color = BAR_BORDER, linewidth = BAR_LW) +
  geom_text(aes(label = count), vjust = -0.4,
            size = 4, fontface = "bold", color = "gray15") +
  scale_fill_manual(values = c("TRUE" = HIGHLIGHT_COLOR, "FALSE" = NON_RESOURCE_COLOR),
                    guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Dice rolls across all games",
       subtitle = "Frequency of each total (7s highlighted)",
       x = "Roll total", y = "Times rolled") +
  theme_colonist()
```

## Victory points over turns

One tab per game. Each line traces a player's VP (settlements, cities, longest road, largest army, and any known dev-card VPs) turn by turn. Hidden VP dev cards that the server keeps private won't show up until they're played or the game ends.

```{r vp-timeline-setup}
vp_named <- d$vp_timeline |>
  left_join(d$rosters |> select(game_id, player_color, username),
            by = c("game_id", "player_color"))

# Colonist in-game color convention: red / blue / green / orange by color index
COLONIST_COLORS <- c(
  "1" = "#d23b3b",
  "2" = "#3b6cb1",
  "3" = "#5aa86a",
  "4" = "#e07a3e"
)

plot_vp_timeline <- function(gid) {
  data <- vp_named |> filter(game_id == gid) |> arrange(turn)
  if (nrow(data) == 0) return(ggplot() + theme_void())
  meta <- d$games |> filter(game_id == gid)
  subtitle_date <- if (nrow(meta) && !is.na(meta$start_time[1])) {
    format(meta$start_time[1], "%b %d, %Y")
  } else ""
  vp_to_win <- if (nrow(meta)) meta$vp_to_win[1] else 10L

  visible_raw <- data |> filter(!is_final)
  final_row   <- data |> filter(is_final)

  # True final VP breakdown per player from endGameState.
  # hidden VP cards = exactly the count of type 2 entries (each = 1 VP).
  # Everything else (settlements, cities, longest road, largest army) is
  # publicly visible and should be on the SOLID step line.
  standings_vp <- d$standings |>
    filter(game_id == gid) |>
    mutate(
      vp_cards = purrr::map_int(vp_breakdown, function(vp) {
        val <- vp[["2"]]
        if (is.null(val)) 0L else as.integer(val)
      }),
      public_final = vp_total - vp_cards
    ) |>
    select(player_color, public_final, vp_cards)

  # Extend the visible step line so it ends at the *true* public final
  # (captures any late longest-road / largest-army / winning-turn builds
  # that happened after the last turn boundary snapshot).
  last_visible_turn <- max(visible_raw$turn)
  public_end <- visible_raw |>
    distinct(game_id, player_color, username) |>
    left_join(standings_vp |> select(player_color, public_final),
              by = "player_color") |>
    mutate(turn = last_visible_turn + 1L,
           vp_total = coalesce(public_final, 0L)) |>
    select(game_id, player_color, username, turn, vp_total)

  visible_data <- bind_rows(visible_raw |> select(game_id, player_color,
                                                  username, turn, vp_total),
                            public_end) |>
    arrange(player_color, turn)

  # Reveal step: dotted from public_final to full weighted VP, shifted a tick to the right
  reveal_turn_x <- last_visible_turn + 2L
  reveal_info <- final_row |>
    left_join(standings_vp, by = "player_color") |>
    mutate(
      turn = reveal_turn_x,
      last_visible_vp = public_final,
      hidden_vp = vp_cards,  # Only VP dev cards are truly hidden
      reveal_label = ifelse(hidden_vp > 0,
                            sprintf("%s (%d VP)  [+%d hidden VP cards]",
                                    username, vp_total, hidden_vp),
                            sprintf("%s (%d VP)", username, vp_total))
    )

  # Turns in the lead (public VP, ties share). Based on visible_raw only
  # (turn-boundary snapshots during play).
  lead_counts <- visible_raw |>
    group_by(turn) |>
    mutate(is_leader = vp_total == max(vp_total)) |>
    ungroup() |>
    group_by(player_color, username) |>
    summarise(turns_led = sum(is_leader), .groups = "drop")

  winner_color <- d$standings |>
    filter(game_id == gid, won) |>
    pull(player_color) |>
    first() %or% NA_integer_

  lead_table <- lead_counts |>
    mutate(is_winner = player_color == winner_color) |>
    arrange(desc(turns_led), desc(is_winner))

  # Axis bounds. Leave headroom at the top for the leaderboard box.
  y_max_data <- max(data$vp_total, vp_to_win)
  y_ceiling <- y_max_data + 4.5
  x_min <- min(visible_raw$turn)
  x_max <- max(visible_raw$turn)
  x_range <- max(x_max - x_min, 1L)

  # Leaderboard box geometry
  box_header_h <- 0.9
  box_row_h    <- 0.95
  n_rows       <- nrow(lead_table)
  box_top      <- y_ceiling - 0.3
  box_bottom   <- box_top - box_header_h - n_rows * box_row_h - 0.4
  box_left     <- x_min - x_range * 0.01
  box_right    <- x_min + x_range * 0.30

  lead_table_pos <- lead_table |>
    mutate(
      row = row_number(),
      y_pos = box_top - box_header_h - (row - 0.5) * box_row_h,
      # W- prefix for winner (renders reliably everywhere, unlike Unicode crowns)
      prefix = ifelse(is_winner, "W - ", "    "),
      label = sprintf("%s%s: %d", prefix, username, turns_led)
    )

  ggplot(visible_data,
         aes(x = turn, y = vp_total,
             color = factor(player_color), group = player_color)) +
    # Dashed "win" line
    geom_hline(yintercept = vp_to_win, linetype = "dashed",
               color = "gray70", linewidth = 0.6) +
    annotate("text", x = x_max, y = vp_to_win + 0.3,
             label = paste0("win = ", vp_to_win, " VP"),
             hjust = 1, color = "gray45", fontface = "bold", size = 3.4) +
    # Grey panel behind the leaderboard
    annotate("rect",
             xmin = box_left, xmax = box_right,
             ymin = box_bottom, ymax = box_top,
             fill = "gray90", color = "gray25", linewidth = 0.9,
             alpha = 0.95) +
    # Leaderboard header
    annotate("text",
             x = (box_left + box_right) / 2, y = box_top - box_header_h / 2,
             label = "Turns in lead",
             hjust = 0.5, vjust = 0.5,
             fontface = "bold", size = 4.3, color = "gray15") +
    # Separator under header
    annotate("segment",
             x = box_left + x_range * 0.015, xend = box_right - x_range * 0.015,
             y = box_top - box_header_h, yend = box_top - box_header_h,
             color = "gray45", linewidth = 0.6) +
    # Leaderboard rows
    geom_text(
      data = lead_table_pos,
      aes(x = box_left + x_range * 0.015, y = y_pos, label = label,
          color = factor(player_color)),
      hjust = 0, fontface = "bold", size = 3.9,
      inherit.aes = FALSE
    ) +
    geom_step(linewidth = 1.5, direction = "hv", alpha = 0.95) +
    # Dotted reveal segment from true public final to full weighted final
    geom_segment(
      data = reveal_info,
      aes(x = last_visible_turn + 1L, xend = turn,
          y = last_visible_vp, yend = vp_total,
          color = factor(player_color)),
      linetype = "dotted", linewidth = 1.2, inherit.aes = FALSE
    ) +
    geom_point(data = reveal_info,
               aes(x = turn, y = vp_total, color = factor(player_color)),
               shape = 21, size = 5, stroke = POINT_STROKE,
               fill = "white", inherit.aes = FALSE) +
    ggrepel::geom_text_repel(
      data = reveal_info,
      aes(x = turn, y = vp_total, color = factor(player_color),
          label = reveal_label),
      hjust = 0, direction = "y", nudge_x = 2.5,
      box.padding = 0.9, point.padding = 0.5,
      force = 5, force_pull = 0.2, max.overlaps = Inf,
      fontface = "bold", size = 3.8,
      segment.color = "gray70", segment.linewidth = 0.5,
      segment.linetype = "dotted", min.segment.length = 0,
      show.legend = FALSE, inherit.aes = FALSE
    ) +
    scale_color_manual(values = COLONIST_COLORS, guide = "none") +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.28)),
                       breaks = pretty_breaks(8)) +
    scale_y_continuous(breaks = seq(0, 15, 2),
                       limits = c(0, y_ceiling)) +
    labs(
      title = paste0("Victory points over turns"),
      subtitle = paste0(subtitle_date,
                        " | dotted tail = hidden VP dev cards revealed at game end"),
      x = "Completed turns", y = "Victory points"
    ) +
    theme_colonist()
}
```

```{r resource-prod-fn}
plot_resource_timeline <- function(gid) {
  data <- d$resource_gains |>
    filter(game_id == gid) |>
    count(game_id, player_color, turn, name = "cards") |>
    left_join(d$rosters |> select(game_id, player_color, username),
              by = c("game_id", "player_color")) |>
    group_by(game_id, player_color, username) |>
    arrange(turn) |>
    mutate(cumulative = cumsum(cards)) |>
    ungroup()

  if (nrow(data) == 0) return(ggplot() + theme_void())

  # Extend each player's line to the game's final turn so they all end together
  max_turn <- max(data$turn, na.rm = TRUE)
  tail_rows <- data |>
    group_by(player_color, username) |>
    slice_max(turn, n = 1, with_ties = FALSE) |>
    ungroup() |>
    filter(turn < max_turn) |>
    mutate(turn = max_turn)
  data <- bind_rows(data, tail_rows) |> arrange(player_color, turn)

  endpoints <- data |>
    group_by(player_color, username) |>
    slice_max(turn, n = 1, with_ties = FALSE) |>
    ungroup()

  # Turns as resource leader — for each turn that has data for any player,
  # anyone tied at the max cumulative total gets credit (ties share).
  turns_seq <- seq_len(max_turn)
  all_players <- endpoints |> select(player_color, username)

  # Carry-forward cumulative for each player across every turn (step line values)
  ff <- all_players |>
    tidyr::crossing(turn = turns_seq) |>
    left_join(data |> select(player_color, turn, cumulative),
              by = c("player_color", "turn")) |>
    group_by(player_color) |>
    arrange(turn) |>
    mutate(cumulative = cumulative |>
             {\(x) { z <- x; for (k in seq_along(z)) if (is.na(z[k])) z[k] <- if (k > 1) z[k-1] else 0L; z }}()) |>
    ungroup()

  lead_counts_r <- ff |>
    group_by(turn) |>
    mutate(is_leader = cumulative == max(cumulative) & cumulative > 0) |>
    ungroup() |>
    group_by(player_color, username) |>
    summarise(turns_led = sum(is_leader), .groups = "drop")

  winner_color <- d$standings |>
    filter(game_id == gid, won) |>
    pull(player_color) |>
    first() %or% NA_integer_

  lead_table <- lead_counts_r |>
    mutate(is_winner = player_color == winner_color) |>
    arrange(desc(turns_led), desc(is_winner))

  # Layout
  y_max_data <- max(data$cumulative, na.rm = TRUE)
  y_ceiling  <- y_max_data + y_max_data * 0.35
  x_min      <- min(data$turn)
  x_max      <- max(data$turn)
  x_range    <- max(x_max - x_min, 1L)

  box_header_h <- y_ceiling * 0.07
  box_row_h    <- y_ceiling * 0.072
  n_rows       <- nrow(lead_table)
  box_top      <- y_ceiling - y_ceiling * 0.02
  box_bottom   <- box_top - box_header_h - n_rows * box_row_h - y_ceiling * 0.025
  box_left     <- x_min - x_range * 0.01
  box_right    <- x_min + x_range * 0.32

  lead_table_pos <- lead_table |>
    mutate(
      row = row_number(),
      y_pos = box_top - box_header_h - (row - 0.5) * box_row_h,
      prefix = ifelse(is_winner, "W - ", "    "),
      label = sprintf("%s%s: %d", prefix, username, turns_led)
    )

  ggplot(data, aes(x = turn, y = cumulative,
                   color = factor(player_color), group = player_color)) +
    geom_step(linewidth = 1.5, direction = "hv", alpha = 0.95) +
    geom_point(data = endpoints,
               shape = 21, size = 5, stroke = POINT_STROKE,
               fill = "white") +
    ggrepel::geom_text_repel(
      data = endpoints,
      aes(label = sprintf("%s (%d)", username, cumulative)),
      hjust = 0, direction = "y", nudge_x = 2.5,
      box.padding = 0.8, point.padding = 0.5,
      force = 5, force_pull = 0.2, max.overlaps = Inf,
      fontface = "bold", size = 3.9,
      segment.color = "gray70", segment.linewidth = 0.5,
      segment.linetype = "dotted", min.segment.length = 0,
      show.legend = FALSE
    ) +
    # Grey panel for "turns as resource leader"
    annotate("rect",
             xmin = box_left, xmax = box_right,
             ymin = box_bottom, ymax = box_top,
             fill = "gray90", color = "gray25", linewidth = 0.9,
             alpha = 0.95) +
    annotate("text",
             x = (box_left + box_right) / 2,
             y = box_top - box_header_h / 2,
             label = "Turns as resource leader",
             hjust = 0.5, vjust = 0.5,
             fontface = "bold", size = 4.1, color = "gray15") +
    annotate("segment",
             x = box_left + x_range * 0.015,
             xend = box_right - x_range * 0.015,
             y = box_top - box_header_h,
             yend = box_top - box_header_h,
             color = "gray45", linewidth = 0.6) +
    geom_text(
      data = lead_table_pos,
      aes(x = box_left + x_range * 0.015, y = y_pos, label = label,
          color = factor(player_color)),
      hjust = 0, fontface = "bold", size = 3.9,
      inherit.aes = FALSE
    ) +
    scale_color_manual(values = COLONIST_COLORS, guide = "none") +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.25)),
                       breaks = pretty_breaks(8)) +
    scale_y_continuous(breaks = pretty_breaks(6),
                       limits = c(0, y_ceiling),
                       expand = expansion(mult = c(0.02, 0))) +
    labs(
      title = "Cumulative resource cards received",
      subtitle = "Cards gained from dice-roll distributions (excludes trades, robber, dev cards)",
      x = "Completed turns", y = "Cumulative cards from rolls"
    ) +
    theme_colonist()
}
```

```{r vp-timeline-tabs, results='asis', fig.height=5.6, fig.width=10.5}
games_ordered <- d$games |> arrange(game_order)
cat("::: {.panel-tabset}\n\n")
for (i in seq_len(nrow(games_ordered))) {
  g <- games_ordered[i, ]
  cat("## Game ", g$game_order, "\n\n", sep = "")
  cat("::: {.panel-tabset}\n\n")
  cat("### VP timeline\n\n")
  print(plot_vp_timeline(g$game_id))
  cat("\n\n### Resource production\n\n")
  print(plot_resource_timeline(g$game_id))
  cat("\n\n:::\n\n")
}
cat(":::\n")
```

## Raw data

Frames in scope as `d$<name>`: `games`, `rosters`, `standings`,
`activity_stats`, `dice_histogram`, `events`, `dice_rolls`, `trade_offers`,
`trades_accepted`, `bank_trades`, `builds`, `resource_gains`, `steals`,
`dev_cards`, `achievements`, `tiles`, `corners`, `settlement_placements`,
`settlement_tiles`, `player_stats`.
